animbook: Visualizing changes in performance measures and demographic affiliations using animation

### 2024-01-11 animbook-journal.Rmd
09/01/2024 To do list with NAs handling

  There was an implementation of the New York Times animation in R (@r) by @Hvitfeldt_2018. This article demonstrates the path generation for each point and how to map them to the animation. Taking inspiration from this example, \CRANpkg{animbook} expands upon the visualization by incorporating additional elements, particularly the proportional shaded area, and introduces new visualizations inspired by the New York Times animation.

  As mentioned about the proportional shaded area, a similar static visualization of the animation is known as the Sankey plot. Sankey plots are flow diagrams that show the movement/change from one state to another, with the width of each arrow representing the proportional changes between states. R packages capable of generating a Sankey plot type include  \CRANpkg{ggalluvial} and \CRANpkg{ggparallel}, both are extensions from the \CRANpkg{ggplot2} {@ggplot2}. The main difference between the Sankey plot visualization and the animation featured in the New York Times lies in their methods of representing the proportion of changes between states. Sankey plots utilize the width of connecting lines, while the New York Times animation employs the density and color of points accompanied by counters. Combining the differences between these two features will be shown in the Visualization design section.

@@

@@

  In the animation, the use of proportional shaded areas serves to enhances the quick comprehension of proportion information, effectively displaying the distribution within each group. This method replaces the coutners seen in the New York Times animation, alleviating information overload for users and reducing the computation required. To represents sub-groups, the plot utilizes color to distinguish differences among observation.

  However, discerning the movement pattern or proportion of sub-groups can be challenging when the visualization lacks a sufficient number of data points. Addressing this, observation interpolation is employed by calculating the sub-group proportion and multiplying it by the desired number of generated points. This approach ensures a more detailed representation of movement patterns, particularly in scenarios where a limited number of data points may impede visual clarity.

  The issue with this type of animation in the New York Times lies in point clustering, particularly when all points start simultaneously. This clustering poses a challenge in distinguishing and tracking individual observations. In the context of time series data (e.g., year), the observations in the animation enter the visualization at different times. It is important to note that this does not necessarily reflect the real-time occurrence of observations. The neglect of the time-dependent nature of observations is a deliberate choice made for visualization purposes, aiming enhance the clarity of the presented information.

  The paths of observations need to be interpolated to create small changes in the position of points to produce the appearance of motion, as mentioned in the Animation tools section. These interpolate functions in \CRANpkg{gganimate} and \CRANpkg{plotly} only generate a linear path between points. However, linear paths are visually less appealing than non-linear paths. Sigmoid curves are commonly used in Sankey diagrams (@Hvitfeldt_2018). However, @Shaffer_2019 argues that a sine curve more accurately represents proportion throughout the path. This is illustrated in Figure \@ref(fig:proportional-shade). The sigmoid accurately represents the proportion at the beginning and end, but as it curves, the shape gets narrower, leading to a less accurate proportion representation than the sine curve.

@@

# Summary

Beginning with inspiration drawn from the New York Times article, this packages provides tools to facilitate the communication of complex data to broader audiences. The paper outlies the anticipated data format for input into the prep function, covering raw, categorized, and animated formats. It delves into the animation concept, persistence of vision, achieved through point interpolation, and the visualizatoin design choices.

The paper further proviedes illustrative examples of software usage through two application to the real world data. Given more time, introducing geatures that allow variation in speeds for each observation could enhance the visualization by introducing an additional dimension. The new animated adn static plots, as demonstrated in the New York Times article, would enable comparisons of demographics at specific points in time. Additionally, the potential re-parameterization of the x-axis from years to ages (e.g., company) could add different details to the visualization.

- positives and negatives about the method
- extensions that would be desirable for the software

>>>>

  @Hvitfeldt_2018 provides R (@r) code to generate an animation similar to that of the New York Times. Taking inspiration from this example, \CRANpkg{animbook} expands upon the visualization in numerous ways and packages the code so that it can be use for a variety of similar animations for studying flow and temporal change data.

  The New York Times visualization is also similar to what is called a Sankey plot. Sankey plots are flow diagrams that show the movement/change from one category levels to another, with the width of each arrow representing the proportional changes between levels. There are several R packages available to generate version of Sankey plots including \CRANpkg{ggalluvial} and \CRANpkg{ggparallel}. The main difference between the Sankey plot visualization and the animation featured in the New York Times lies in their methods of representing the proportion of changes. Sankey plots utilize the width of connecting lines, while the New York Times animation employs the density and color of points.

@@

  One of the frequently encountered values in the dataset is NA, commonly known as missing values. In the context of accounting data, NA might signify that a company is either not yet listed or has been de-listed. In scenarios like voter data, NA may indicate that individuals did not provide the necessary information. Disregarding NA values could result in the loss of valuable information. Therefore, instead of completely removing NA entries, this information is categorized into its quantiles, ensuring that the missing data are factored into the visualization.

@@

  In the animation, the use of proportional shaded areas serves to enhances the comprehension of proportion information, effectively displaying the distribution within each group. To represent sub-groups, points are colored to focus attention on differences between groups of observation.

  Discerning the movement pattern or proportion of sub-groups can be challenging when there is only a small number of points. To address this, sample size are inflated by calculating the sub-group proportion and multiplying it by the desired total number of points. This approach ensures a richer representation of movement patterns.

 Another issue is that all observations have a common starting time. This would mean all points start the animation simultaneously and move as a block through time. This poses a challenge in distinguishing and tracking individual observations and their change patterns. To alleviate this, the animation randomly starts observations at different times so a continuous flow of points is shown. It is important to note that this does not necessarily reflect the real-time occurrence of observations.

  The paths of observations are also interpolated to create small changes in the position of points to produce the appearance of motion. The interpolation functions in \CRANpkg{gganimate} and \CRANpkg{plotly} are inadequate because they only generate a linear path. Our animation needs non-linear path to show the flow. Sigmoid curves are commonly used in Sankey diagrams (@Hvitfeldt_2018). However, @Shaffer_2019 argues that a sine curve more accurately represents proportion throughout the path. As illustrated in Figure \@ref(fig:proportional-shade). The sigmoid accurately represents the proportion at the beginning and end, but as it curves, the shape gets narrower, leading to a less accurate proportion representation than the sine curve. New interpolation functions are needed to compute both the sigmoid and sine path.

@@

# Discussion

Beginning with inspiration drawn from the New York Times article, this packages provides tools to facilitate the communication of complex data to broader audiences. The paper outlines the anticipated data format for input into the prep function, covering raw, categorized, and animated data formats. It delves into the animation concept, persistence of vision, achieved through point interpolation, and the visualization design choices.

  The \CRANpkg{animbook} package comes with certain drawbacks. Firstly, it leans more towards being a communication tool rather than an exploratory one, as the information it presents can often be obtained from a summary table. Secondly, it may struggle to handle datasets with numerous categories (quantiles), limiting its applicability in such scenarios. Furthermore, perceiving patterns from the animation might be challenging for users when there are differences in sample sizes across groups. Lastly, the animation requires the reader to mentally calculate the fraction of points that are in motion. However, the package provides tools that provides a simple solution for creating meaningful animations through a simple step process.

  Introducing features that allow variation in speeds for each observation could enhance the visualization by introducing an additonal dimension. The new animated and static plots, as demonstrated in the New York Times article, would enable comparisons of demographics at specific points in time. Additionally, the potential re-parameterization of the x-axis from years to ages (e.g., company ages) could add different ways of viewing the visualization.

- the first paragraph is to summarize the achievement of this
- the second paragraph is about the pro and con, what works, what doesn't works.
- the third paragraph, if you have another 6 months, what would you implemented.

### 2024-01-11 animbook-journal.tex
09/01/2024 To do list with NAs handling

There was an implementation of the New York Times animation in R (R Core Team (2021)) by Hvitfeldt (2018). This article demonstrates the path generation for each point and how to map them to the animation. Taking inspiration from this example, \CRANpkg{animbook} expands upon the visualization by incorporating additional elements, particularly the proportional shaded area, and introduces new visualizations inspired by the New York Times animation.

As mentioned about the proportional shaded area, a similar static visualization of the animation is known as the Sankey plot. Sankey plots are flow diagrams that show the movement/change from one state to another, with the width of each arrow representing the proportional changes between states. R packages capable of generating a Sankey plot type include \CRANpkg{ggalluvial} and \CRANpkg{ggparallel}, both are extensions from the \CRANpkg{ggplot2} \{Wickham (2016)\}. The main difference between the Sankey plot visualization and the animation featured in the New York Times lies in their methods of representing the proportion of changes between states. Sankey plots utilize the width of connecting lines, while the New York Times animation employs the density and color of points accompanied by counters. Combining the differences between these two features will be shown in the Visualization design section.

@@

In the animation, the use of proportional shaded areas serves to enhances the quick comprehension of proportion information, effectively displaying the distribution within each group. This method replaces the coutners seen in the New York Times animation, alleviating information overload for users and reducing the computation required. To represents sub-groups, the plot utilizes color to distinguish differences among observation.

However, discerning the movement pattern or proportion of sub-groups can be challenging when the visualization lacks a sufficient number of data points. Addressing this, observation interpolation is employed by calculating the sub-group proportion and multiplying it by the desired number of generated points. This approach ensures a more detailed representation of movement patterns, particularly in scenarios where a limited number of data points may impede visual clarity.

The issue with this type of animation in the New York Times lies in point clustering, particularly when all points start simultaneously. This clustering poses a challenge in distinguishing and tracking individual observations. In the context of time series data (e.g., year), the observations in the animation enter the visualization at different times. It is important to note that this does not necessarily reflect the real-time occurrence of observations. The neglect of the time-dependent nature of observations is a deliberate choice made for visualization purposes, aiming enhance the clarity of the presented information.

The paths of observations need to be interpolated to create small changes in the position of points to produce the appearance of motion, as mentioned in the Animation tools section. These interpolate functions in \CRANpkg{gganimate} and \CRANpkg{plotly} only generate a linear path between points. However, linear paths are visually less appealing than non-linear paths. Sigmoid curves are commonly used in Sankey diagrams (Hvitfeldt (2018)). However, Shaffer (2019) argues that a sine curve more accurately represents proportion throughout the path. This is illustrated in Figure \ref{fig:proportional-shade}. The sigmoid accurately represents the proportion at the beginning and end, but as it curves, the shape gets narrower, leading to a less accurate proportion representation than the sine curve.

@@

\hypertarget{summary}{%
\section{Summary}\label{summary}}

Beginning with inspiration drawn from the New York Times article, this packages provides tools to facilitate the communication of complex data to broader audiences. The paper outlies the anticipated data format for input into the prep function, covering raw, categorized, and animated formats. It delves into the animation concept, persistence of vision, achieved through point interpolation, and the visualizatoin design choices.

The paper further proviedes illustrative examples of software usage through two application to the real world data. Given more time, introducing geatures that allow variation in speeds for each observation could enhance the visualization by introducing an additional dimension. The new animated adn static plots, as demonstrated in the New York Times article, would enable comparisons of demographics at specific points in time. Additionally, the potential re-parameterization of the x-axis from years to ages (e.g., company) could add different details to the visualization.

\begin{itemize}
\tightlist
\item
  positives and negatives about the method
\item
  extensions that would be desirable for the software
\end{itemize}

@@

\leavevmode\vadjust pre{\hypertarget{ref-ggplot2}{}}%
---------. 2016. \emph{Ggplot2: Elegant Graphics for Data Analysis}. Springer-Verlag New York. \url{https://ggplot2.tidyverse.org}.

>>>>

Hvitfeldt (2018) provides R (R Core Team (2021)) code to generate an animation similar to that of the New York Times. Taking inspiration from this example, \CRANpkg{animbook} expands upon the visualization in numerous ways and packages the code so that it can be use for a variety of similar animations for studying flow and temporal change data.

The New York Times visualization is also similar to what is called a Sankey plot. Sankey plots are flow diagrams that show the movement/change from one category levels to another, with the width of each arrow representing the proportional changes between levels. There are several R packages available to generate version of Sankey plots including \CRANpkg{ggalluvial} and \CRANpkg{ggparallel}. The main difference between the Sankey plot visualization and the animation featured in the New York Times lies in their methods of representing the proportion of changes. Sankey plots utilize the width of connecting lines, while the New York Times animation employs the density and color of points.

@@

In the animation, the use of proportional shaded areas serves to enhances the comprehension of proportion information, effectively displaying the distribution within each group. To represent sub-groups, points are colored to focus attention on differences between groups of observation.

Discerning the movement pattern or proportion of sub-groups can be challenging when there is only a small number of points. To address this, sample size are inflated by calculating the sub-group proportion and multiplying it by the desired total number of points. This approach ensures a richer representation of movement patterns.

Another issue is that all observations have a common starting time. This would mean all points start the animation simultaneously and move as a block through time. This poses a challenge in distinguishing and tracking individual observations and their change patterns. To alleviate this, the animation randomly starts observations at different times so a continuous flow of points is shown. It is important to note that this does not necessarily reflect the real-time occurrence of observations.

The paths of observations are also interpolated to create small changes in the position of points to produce the appearance of motion. The interpolation functions in \CRANpkg{gganimate} and \CRANpkg{plotly} are inadequate because they only generate a linear path. Our animation needs non-linear path to show the flow. Sigmoid curves are commonly used in Sankey diagrams (Hvitfeldt (2018)). However, Shaffer (2019) argues that a sine curve more accurately represents proportion throughout the path. As illustrated in Figure \ref{fig:proportional-shade}. The sigmoid accurately represents the proportion at the beginning and end, but as it curves, the shape gets narrower, leading to a less accurate proportion representation than the sine curve. New interpolation functions are needed to compute both the sigmoid and sine path.

@@

\hypertarget{discussion}{%
\section{Discussion}\label{discussion}}

Beginning with inspiration drawn from the New York Times article, this packages provides tools to facilitate the communication of complex data to broader audiences. The paper outlines the anticipated data format for input into the prep function, covering raw, categorized, and animated data formats. It delves into the animation concept, persistence of vision, achieved through point interpolation, and the visualization design choices.

The \CRANpkg{animbook} package comes with certain drawbacks. Firstly, it leans more towards being a communication tool rather than an exploratory one, as the information it presents can often be obtained from a summary table. Secondly, it may struggle to handle datasets with numerous categories (quantiles), limiting its applicability in such scenarios. Furthermore, perceiving patterns from the animation might be challenging for users when there are differences in sample sizes across groups. Lastly, the animation requires the reader to mentally calculate the fraction of points that are in motion. However, the package provides tools that provides a simple solution for creating meaningful animations through a simple step process.

Introducing features that allow variation in speeds for each observation could enhance the visualization by introducing an additonal dimension. The new animated and static plots, as demonstrated in the New York Times article, would enable comparisons of demographics at specific points in time. Additionally, the potential re-parameterization of the x-axis from years to ages (e.g., company ages) could add different ways of viewing the visualization.

\begin{itemize}
\tightlist
\item
  the first paragraph is to summarize the achievement of this
\item
  the second paragraph is about the pro and con, what works, what doesn't works.
\item
  the third paragraph, if you have another 6 months, what would you implemented.
\end{itemize}

@@
